---
title: "Predicted Butterfly Emergence - Visualization"
author: "Elsita Kiekebusch"
date: "9/11/2019"
#output: pdf_document
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
require(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Extracting First Emergence Date  

Random Forest has predicted daily butterfly presence and absence for 20 General Circulation Models (GCMs) paramaterized by the RCP 8.5 emissions scenario (aka the "business as usual scenario"). These are collected in the file "Output.All.PresenceAbsences.RCP8.5.csv" with each column representing 1 GCM. Here we extract the first date of butterfly emergence for each year.


```{r predictions.all.GCMs, warning=FALSE}
setwd('~/Visualizing-Butterfly-Emergence')

allData <- read.csv(file="Output.ALL.PresenceAbsence.RCP8.5.csv")

#get years
emerge <- as.data.frame(unique(allData$Year))
names(emerge)[1] <- "Year"

for (i in 3:22){
  #remove all absence data for each column of interest
  presence <- allData[!(allData[i] == "1:ABSENT"),] 
  ###extracting earliest emergence dates
  calc <- presence %>%
    group_by(Year) %>%
    summarise(early = min(Ordinal))
  emerge <- cbind(emerge,calc$early)
}
```

```{r name columns, include=FALSE}
#rename columns (for each GCM)
names(emerge)[2] <- "bcc-csm1-1_rcp45"
names(emerge)[3] <- "bcc-csm1-1-m_rcp45"   
names(emerge)[4] <- "BNU-ESM_rcp45"
names(emerge)[5] <- "CanESM2_rcp45"
names(emerge)[6] <- "CCSM4_rcp45"
names(emerge)[7] <- "CNRM-CM5_rcp45"
names(emerge)[8] <- "CSIRO-Mk3-6-0_rcp45"
names(emerge)[9] <- "GFDL-ESM2G_rcp45"
names(emerge)[10] <- "GFDL-ESM2M_rcp45"
names(emerge)[11] <- "HadGEM2-CC365_rcp45"
names(emerge)[12] <- "HadGEM2-ES365_rcp45"
names(emerge)[13] <- "inmcm4_rcp45"
names(emerge)[14] <- "IPSL-CM5A-LR_rcp45"
names(emerge)[15] <- "IPSL-CM5A-MR_rcp45"
names(emerge)[16] <- "IPSL-CM5B-LR_rcp45"
names(emerge)[17] <- "MIROC-ESM_rcp45"
names(emerge)[18] <- "MIROC-ESM-CHEM_rcp45"
names(emerge)[19] <- "MIROC5_rcp45"
names(emerge)[20] <- "MRI-CGCM3_rcp45"
names(emerge)[21] <- "NorESM1-M_rcp45"
```


```{r look}
#look at the data real quick
str(emerge)
```


## Visualize the Emergence Dates

Create a plot for each GCM to have a first look at the data. Generally, (and in accordance with the literature), butterflies are emerging earlier and earlier every year as temperatures rise due to climate change. 


```{r viz1}
#tidy, multiple plots
emerge %>% tidyr::gather("id", "Ordinal.Date", 2:21) %>% 
  ggplot(., aes(Year, Ordinal.Date))+
  geom_line()+
  #geom_point()+ #makes a point not a line
  #geom_smooth(method = "lm", se=FALSE, color="black")+ #can give you a linear regression line for each panel
  facet_wrap(~id)
```

## Mean Emergence

Calculate the mean emergence date across the 20 GCMs as well as 95% Confidence Intervals, and display the plot. Confidence intervals can be displayed as separate lines (first figure) or as shading (second figure).


```{r viz2}
#get mean
emerge$Mean <- NA
emerge$Mean <- rowMeans(emerge[,2:21])

#get quantiles
emerge$Upper <- NA
emerge$Lower <- NA

#formulas
qUp = function(x) quantile(x,0.975)
qLo = function(x) quantile(x,0.025)

#upper and lower 95% confidence intervals for all values
emerge$Upper <- apply(emerge[,2:21],1,qUp)
emerge$Lower <- apply(emerge[,2:21],1,qLo)

#plot
ggplot(emerge, aes(Year, Mean)) +
geom_line() +
geom_line(aes(y = Upper, color = "red")) +             
geom_line(aes(y = Lower, color = "red")) + 
theme(legend.position="none") + 
labs(y="Ordinal Date of First Emergence")

#another way to visualize
ggplot(emerge, aes(Year)) +
  labs(x="Year", y="Ordinal Date of First Emergence") +
  geom_line(aes(y=Mean)) +
  geom_ribbon(aes(ymin = Upper, ymax=Lower), alpha=0.1) +
  theme_classic()

```
